Implementing Real-time Transport Services over an Ossified Network

### 2016-05-16 paper.tex
Work on Innovation and Ossification

@@

Briefly discuss: the Internet architecture, in principle, allows free
innovation at the transport layer, provided the IP layer is unchanged. 
Outline why this is not true in practice.

Implication: new protocols must look like TCP or UDP.

UDP is the obvious choice, since it provides minimal additional services
over the IP layer, allowing great flexibility in innovation for protocols
tunnelled on top. Only real cost is a few bytes of additional header.

But, UDP is often blocked. Corporate firewalls. History of abuse for DDoS
attacks (draft-byrne-opsec-udp-advisory-00). Slowly changing, e.g., due to
QUIC, WebRTC, but not clear that UDP is universally available.

TCP is a more complex choice for innovation. More sophisticated protocol.
Complex headers. Mandates much more behaviour.

However:
- end-points can change congestion control, provided it's ACK-based
- segments need be sent with contiguous sequence number space, resending
  segments to fill gaps in sequence number space caused by packet loss
- traditional API is restrictive, but flexibility to do out-of-order 
  delivery to applications
- easy to perform consistent segmentation on sending, and add COBS framing, 
  to get message-oriented semantics

>>>>

\begin{figure}
  \centering
  ...
  \caption{Outline transport API for real-time applications}
  \label{fig:api}
\end{figure}

@@

The Internet architecture, in principle, allows free innovation at the
transport layer, provided the underlying network (IP) layer is unchanged.  
Routers should inspect the source addresses of packets to perform network
ingress filtering \cite{RFC2827}, and the destination addresses to route
packets to the correction destination, but should not inspect their
contents. This is not, of course, how the real network operates.

\todo{outline why this is not true in practice}

The implication of this reality is that it is difficult to deploy new 
transport protocols. The installed base of NATs, firewalls, and other
middleboxes is such that packets that do not look like TCP or UDP are
unlikely to pass the network. We may innovate all we like, provided the
transport of the future looks like TCP or UDP to middleboxes.\footnote{
  This is inconvenient, certainly, but is not necessarily a bad thing. 
  The Internet is critical infrastructure. It support emergency services, 
  healthcare applications, infrastructure components, financial services, 
  and so on, many of which are essential to the functioning of society. 
  Making changes to this type of infrastructure \emph{should} require 
  careful backwards compatibility.
}

UDP is the obvious base for future protocol development, since it provides
minimal additional services over the IP layer, allowing great flexibility
in innovation for protocols tunnelled on top. Provided middleboxes do not
inspect the payload too carefully, the only real cost to innovation, when
compared to a native transport protocol running over IP, is a few bytes of
additional header. Examples in this space include RTP \cite{RFC3550}, one
of the most widely deployed real-time transport protocols; the WebRTC Data
Channel \cite{}, which tunnels peer-to-peer SCTP associations over a DTLS
association over UDP; and QUIC \cite{}, which provides a modern alternative 
to TCP, implemented over UDP.

Despite these advantages, UDP can be problematic as a substrate for new
protocol development. UDP traffic is blocked by some enterprise firewalls,
and some in the operations community have a strong distrust of UDP-based
protocols and applications \cite{draft-byrne-opsec-udp-advisory-00}. In
part this is due to ignorance. Outside specific niches, such as DNS, UDP
has not been widely used in enterprise environments, and hence is widely
misunderstood. Blocking the unknown is a rational response. In addition,
UDP traffic has been widely used as a component of distributed denial of
service (DDoS) attacks, leading some to install blanket blocks of UDP as 
a safety measure (blanket blocking, rather than the more targeted blocks 
used when TCP traffic is used in DDoS attacks, are justified using the
argument that UDP is not widely used).
These issues are slowly changing, as UDP-based applications penetrate the
enterprise consciousness, but not clear that UDP is universally available
(Google report XX\% \todo{how much?} of endpoints are reachable with QUIC
running over UDP \cite{}).

TCP is a more complex choice for innovation. It is a more sophisticated 
protocol than UDP, with complex headers, and a protocol state machine
that mandates much more behaviour and is widely understood, and policed,
by in-network middleboxes. This does not mean that TCP cannot evolve, or
form the basis for new transport services. Rather, it means that any
innovation or development must be done carefully, paying very careful
attention to backwards compatibility.

We identify a number of places where TCP can evolve with comparative
freedom. 
Firstly, the TCP congestion control algorithm is executed by the end
points, and can be changed provided the new version requires no new
information to be exchanged. 
\todo{expand, and discuss how this evolution has occurred to date}

Then, we note that the API that is exposed to applications using TCP is
not visible to the network. 
- traditional API is restrictive, but flexibility to do out-of-order 
  delivery to applications

- easy to perform consistent segmentation on sending, and add COBS framing, 
  to get message-oriented semantics

- segments need be sent with contiguous sequence number space, resending
  segments to fill gaps in sequence number space caused by packet loss
